Comb associative operads

### 2018-06-29 Magmatic_operads.tex
Rédaction première partie opérades magmatiques

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Magmatic operads}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Category of the magmatic quotients}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A lattice of magmatic quotients}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Magmatic quotients and combinatorics}

% Hilbert series, enumeration, a little of rewrite systems, etc.

>>>>

\section{Quotients of the magmatic operad}

\Todo{Write mini intro.}

\subsection{Lattice structure}

We denote by $\IMag$ the set of operadic of $\Mag$, and
\begin{equation} \label{equ:definition_of_QMag}
  \QMag =\Big\{\Mag/ I\mid I\in\IMag\Big\},
  \end{equation}
the set of quotients of the magmatic operad. Given two elements $\Oca_1$
and $\Oca_2$ of $\QMag$, we denote by $\text{hom}\left(\Oca_1,\Oca_2\right)$
the set of operads morphisms from $\Oca_1$ to $\Oca_2$.

\begin{Proposition} \label{prop:endomorphisms_of_magmatic_operads}
  The set $\emph{hom}\left(\Oca_1,\Oca_2\right)$ admits a vector space
  structure. Moreover, its dimension is equal to $0$ or $1$ and every
  nonzero morphism is surjective.
\end{Proposition}

\begin{proof}
  The operad $\Oca_1$ is spanned by the binary generator~$\Product$, so
  that a morphism $\varphi:\Oca_1\to\Oca_2$ is uniquely determined by
  $\varphi(\Product)$. Moreover, $\varphi(\Product)$ has arity $2$ is
  $\Oca_2$. Hence, $\varphi(\Product)$ belongs to the line spanned by
  the binary generator of $\Oca_2$, that is there exists a scalar
  $\lambda\in\K$ such that $\varphi(\Product)=\lambda\Product$. If there
  exists such a $\lambda$ different from zero, then for every nonzero
  scalar $\mu$, we have a well-defined operads morphism
  $\psi:\Oca_1\to\Oca_2$ satisfying $\psi(\Product)=\mu\Product=\left(
  \mu/\lambda\right)\varphi(\Product)$. Hence,
  $\text{hom}\left(\Oca_1,\Oca_2\right)$ is either reduced to the zero
  morphism or it is in one-to-one correspondance with $\K$, which
  proves that $\text{hom}\left(\Oca_1,\Oca_2\right)$ is a vector space
  of dimension at most $1$. Moreover, if $\varphi$ is different from $0$,
  that is there is a nonzero scalar such that
  $\varphi(\Product)=\lambda\Product$, we have
  $\varphi(1/\lambda\Product)=\Product$, so that $\varphi$ is
  surjective.
\end{proof}

We introduce the binary relation $\preceq$ on $\QMag$ as follows: we
have $\Oca_2\preceq\Oca_1$ if the dimension of
$\text{hom}\left(\Oca_1,\Oca_2\right)$ is equal to $1$.

\begin{Proposition} \label{prop:order_relations_on_QMag_and_ideals}
  Let $\Oca_1=\Mag/I_1$ and $\Oca_2=\Mag/I_2$ be two elements of $\QMag$.
  We have $\Oca_2\preceq\Oca_1$ if and only if $I_1\subseteq I_2$.
\end{Proposition}

\begin{proof}
  Since $\text{hom}\left(\Oca_1,\Oca_2\right)$ is a vector space of
  dimension at most $1$, it contains a nonzero morphism if and only if
  the morphism
  $\overline{\varphi}:\Oca_1\to\Oca_2$ satisfying
  $\overline{\varphi}(\Product)=\Product$ is well-defined, which means
  that $\Oca_2\preceq\Oca_1$ is equivalent to this condition. Moreover,
  by the universal property of the quotient, $\overline{\varphi}$ is
  well-defined if and only if $I_1$ is included in the kernel of the
  morphism $\varphi:\Mag\to\Oca_2$ defined by
  $\varphi(\Product)=\Product$. This kernel is equal to $I_2$, so that
  $\overline{\varphi}$ is well-defined if and only if $I_1$ is included
  in $I_2$, which concludes the proof.
\end{proof}

Recall that a \Def{lattice} is a tuple $\left(E,\leq,\wedge,\vee\right)$,
where $\left(E,\leq\right)$ is a poset such that any two elements $e$
and $e'$ of $E$ admit a lower-bound $e\wedge e'$ and an upper-bound
$e\vee e'$. In particular, $\left(\IMag,\subseteq,\cap,+\right)$ is a
lattice, where $\cap$ and $+$ are the intersection and the sum of
operadic ideals, respectively.

Given two elements $\Oca_1=\Mag/I_1$ and $\Oca_2=\Mag/I_2$ of $\QMag$,
we let
\begin{equation} \label{equ:definition_of_lattice_operations_of_QMag}
  \Oca_1\wedge \Oca_2=\Mag/\left(I_1+I_2\right)\ \ \text{and}\ \
  \Oca_1\wedge \Oca_2=\Mag/\left(I_1\cap I_2\right).
  \end{equation}

\begin{Theorem} \label{thm:lattice_structure_of_QMag}
  The tuple $\left(\QMag,\preceq,\wedge,\vee\right)$ is a lattice.
\end{Theorem}

\begin{proof}
  The map $\mathcal{Q}:\IMag\to\QMag$ defined by $\mathcal{Q}(I)=\Mag/I$
  is a bijection by definition of $\QMag$. Moreover, from
  Proposition~\ref{prop:order_relations_on_QMag_and_ideals},
  $\Oca_2\preceq\Oca_1$ is equivalent to $I_1\subseteq I_2$, so that
  $\left(\QMag,\preceq\right)$ is a poset and $\mathcal{Q}$ is a
  decreasing bijection. The tuple $\left(\IMag,\subseteq,\cap,+\right)$
  being a lattice, the decreasing bijection $\mathcal{Q}$ induces
  lattice operations on $\QMag$, precisely $\wedge$ and $\vee$ by
  definition.
  \end{proof}
\subsection{Kernels of magmatic operads}
